### 2013-02-12 supplementary.tex
Fig.1 has p10; Fig S3 -> S3/S4; more details on epitopes; S2 larger

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template for a revtex article
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[rmp]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,eucal,eurosym,textcomp}
\usepackage{color}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage{natbib}
\usepackage{pslatex}
\usepackage[colorlinks,linkcolor=red,citecolor=red]{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./figures/}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\comment}[1]{{\color{red}#1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\thesubfigure}{\Alph{subfigure}}
\newcommand{\Author}{Fabio~Zanini* and Richard~A.~Neher*}
\newcommand{\Title}{Deleterious synonymous mutations hitchhike to high frequency in HIV \env~evolution}
\newcommand{\Affiliation}{Max Planck Institute for Developmental Biology, 72076 T\"ubingen, Germany}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\Title}
\author{\Author}
\affiliation{\textsf{*}\Affiliation}
\date{\today}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Selection of the patient data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\linewidth]{Shankarappa_PCA_p1}
\includegraphics[width=0.35\linewidth]{Shankarappa_allele_freqs_trajectories_nonsyn_p1}
\includegraphics[width=0.35\linewidth]{Shankarappa_PCA_p7}
\includegraphics[width=0.35\linewidth]{Shankarappa_allele_freqs_trajectories_nonsyn_p7}
\caption{Structure of viral population and patient selection.
Panel A) shows a PCA of all sequences from patient p1 (colors indicate time from
seroconversion, from blue to red). Panel B) Allele frequency trajectories for nonsynonymous
changes in the same patient. Panels C and D) show analogous plots for data from
patient p7. Samples after day 1000 split into two clusters in the PCA and no
mutations that arise after day 1000 fix, presumably because they are restricted
to one subpopulation. All patients like p7 (p4, p7, p8, p9 from ref.~\citealp{shankarappa_consistent_1999} and
ACH19542 and ACH19768 from ref.~\citealp{bunnik_autologous_2008}) were excluded
from our analysis.}
\label{fig:aftp}
\end{center}
\end{figure}

\newpage
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synonymous diversity across the HIV genome}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
\begin{center}
\includegraphics[width=\linewidth]{conservation_codons_genome}
\caption{
Synonymous diversity across the HIV genome, as quantified by the normalized
codon entropy among sequences coding for the consensus amino-acid. The
synonymous diversity peaks at the variable regions in {\it env} and is reduced
in regions under purifying selection (RRE hairpin, second {\it tat}/{\it rev}
exons). Since the variable regions are under positive selection at the protein
level, synonymous hitchhiking is stronger there, and even deleterious synonymous
mutations can be brought to high frequency. This gives to our signal in the
fixation probability, Figs. 2A and 2B of the main text. The normalized codon
entropy is calculated as follows (see the script for the full algorithm): (i)
from a subtype B multiple sequence alignment (MSA) from the LANL website
(filtered sequences only, version 2011)~\cite{LANL2012}, we calculate the
consensus amino acid at each position in the HIV genome; (ii) we count how often
each codon coding for the consensus amino acid appears in the MSA; (iii) at each
amino acid position, we divide by the number of sequences in the MSA that had
the consensus amino acid at that position, obtaining {\it codon frequencies}
$\nu_c$; (iv) we calculate the codon entropy from each position as: $S := -
\sum_{c} \nu_c \log \nu_c$, where $c$ runs over codons that code for the
consensus amino acid at this site; (v) we divide by the maximal codon entropy of
that amino acid (e.g. $\log 2$ for twofold degenerate codons). All parts of
{\it env} that are part of a different gene (signaling peptide, second {\it rev}
exon) have been excluded from our main analysis, to avoid contamination about
synonymity. Note that all gap-rich columns of the MSA are stripped from this
figure, hence genes such as {\it env} might appear shorter than they actually
are.
}
\label{fig:syndiv_genome}
\end{center}
\end{figure}
\newpage
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Nonsynonymous changes outside of variable regions are deleterious}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=0.8\linewidth]{synmut_conservation_4fold_synnonsyn}
% \caption{Cumulative distribution of synonymous and nonsynonymous diversity in
% {\it gag} in the LANL reference panel (filtered
% sequences only, version 2011)~\cite{LANL2012}. Sites such that the consensus codon
% has three synonymous and six nonsynonymous single mutants were used, and the number of observed mutants
% of a certain type over the number of possible mutants is plotted. Nonsynonymous
% changes are observed less often, {\it ergo} are more conserved, than
% synonymous changes. It can be therefore assumed, as mentioned in the main text,
% that non-escape nonsynonymous changes involve a large fitness cost.}
% \label{fig:synnonsyncons}
% \end{center}
% \end{figure}
% \newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Time-dependent selection}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\linewidth]{simulations_gradually_timeselec_deadagain}
\caption{
Time-dependent selection (the escape mutant is recognized after some months from
its appearance) reduces fixation of nonsynonymous
mutations. A ``na\"ive'' computer simulation shows a depression in synonymous
fixation probability, as observed in patients, but the nonsynonymous fixation
probability is much higher than the (neutral) diagonal: a nonsynonymous mutation
that reaches a frequency around 0.2 is already very likely to fix. In data from
infected patients, however, nonsynonymous mutations show a neutral-like curve,
$P_\text{fix}(\nu) \approx \nu$. In the other curves, we show that the
nonsynonymous fixation probability can be reduced by time-dependent selection
without affecting much the synonymous fixation probability itself. The fitness
advantage of an escape mutaition can vanish at each time point with a
probability density \[ P_\text{recognized}(t) = c \cdot \nu(t), \] where $c$ is
a constant coefficient shown in the legend that encodes the overall efficiency
of the host immune system, and $\nu(t)$ is the frequency of the escape allele at
time $t$. After recognition, the former escape mutation has the same fitness
cost as before the creation of the epitope, hence it is quickly lost. Note that,
although the curves shown here show almost no reduction in fixation probability
up to $\nu \approx 0.15$, this effect is not robust and depends on the exact form of
$ P_\text{recognized}(t)$. The other parameters for the shown simulations are
the following: deleterious effect $10^{-3}$, average escape rate $0.016$,
fraction of deleterious synonymous mutations $0.986$, rate of new epitopes
$0.0014$ per generation.
}
\label{fig:tds}
\end{center}
\end{figure}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Within-epitope competition}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\linewidth]{simulations_gradually_epitopes}
\caption{
Competition between escape mutations in the same epitope also reduces fixation of
nonsynonymous mutations. Compared to a ``na\"ive'' model,
simulations with 3 or 6 escape mutations competing within the same epitope show
reduced fixation probabilities of nonsynonymous changes, whereas the synonymous
changes behave similarly in all cases. The epitope length is 9 nucleotides. The
fitness landscape of each such epitope include negative epistatic terms, so that
the joint presence of more than one escape mutation is not any more beneficial
for the virus than a single mutation. In detail, each site has two alleles, $\pm
1$, where $-1$ is the ancestral one and $+1$ the dervied one; the fitness
coefficient of a $k$-uple of sites within the epitope is $f_k = \pm
\eta_\epsilon / 2^{L-1}$, where $\eta_\epsilon$ is a random variable for the
escape rate with average value $\epsilon$ (see methods), $L$ is the number of
competing escapes in the epitope, and the positive (resp. negative) sign is for
odd (resp. even) $k$. In this evolutionary scenario, many escape mutations start
to sweep on different backgrounds within the viral population, but eventually
compete and only one of them if fixed. The other parameters for the shown simulations are
the following: deleterious effect $10^{-3}$, average escape rate $0.016$,
fraction of deleterious synonymous mutations $0.986$, rate of new epitopes
$0.0014$ per generation.
}
\label{fig:wec}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{natbib}
\bibliography{bib}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%