### 2013-08-17 response_to_reviewers.txt
minor edits

##############################################################################
# Reviewer comments:
##############################################################################
# Reviewer #1 (Comments for Author):
##############################################################################

>The text is clearly written and the written arguments are well spelled out.
>They observe that synonymous mutations in the C2-V5 region that disrupt
>modelled base pairs in SHAPE assays are more likely to be lost than other
>mutations, implying clear fitness costs to the virus, although they point out
>that those with major costs will never reach high frequencies. They argue that
>less deleterious mutations, but still with a fitness cost, can rise to high
>frequency by 'hitch hiking' with immune selected non-synonymous mutations, in
>other words simple linkage. The selection against synonymous slows down
>evolution of the virus so selective forces may have been underestimated in the
>past. These are important concepts and show that immune escape mutations
>should not be considered in isolation. They support the view that 'landscaping'
>approaches such as that described recently by Chakraborty in Immunity (Immunity
>38:606-617, 2013), need to be applied.

We thank the reviewer for this thoughtful assessment. We have included a
brief discussion of the landscape aspect of HIV fitness and now cite
Ferguson et al, Immunity, 2013.

##############################################################################
# Reviewer #2 (Comments for Author):
##############################################################################

>These are interesting results and the manuscript is clearly written. The second
>observation is not novel, but underappreciated, whereas this is the first
>concrete calculation of the average selective disadvantage of silent mutations
>in HIV taking draft into account. As such, it is necessary for this work to
>establish proper error estimates so that the community can make valid use of
>the results.  With this in mind, the manuscript can be strengthened in the
>following respects:

We thank the reviewer for the constructive comments, which have helped
us to improve the manuscript. We have resorted to a more sophisticated
sampling scheme for parameters in the computer simulations and now scan
a larger parameter space. All our conclusions remain valid, but are based
on more exhaustive sampling. We detail these improvements in our point to point
response below.

##############################################################################
>1) The modeling generation time of 1 day is odd.  It takes about a day for the
>virion to integrate and start transcription.  One usually takes 2 days as the
>mean generation time, but there is individual variation. See next point.

We agree that one generation in HIV takes closer to 2 days than 1
day. Most escape rates and selection coefficients, however, are more
readily measured in days and we have chosen to model HIV evolution with
generation of one day for simplicity. Factor of 2 changes in the
generation have little influence on the dynamics and only renormalize
genetic drift. Genetic drift is not a dominant force in evolution of the
HIV envelope gene. The (effective) population size, the main determinant of drift,
is not known to this accuracy anyway. We have repeated a subset of the
simulations with a generation time of one day, finding no
difference. Our more exhaustive survey of the parameter space includes a
10-fold variation of the population size. Our results are very
insensitive and hence also insensitive to the generation time.

##############################################################################
>2) Related to the above: the simulation does not address how sensitive the
>results are to varying the parameters (in particular, the generation time, the
>outcrossing rate, the mutation rate, and the assigned deleterious fitness of
>the "non-synonymous" changes).  In the absence of described theoretical
>analysis shaping our expectations, I strongly feel that this sensitivity
>analysis needs to be part of the results, and folded into the error bars.

We have greatly extended the parameter space that we survey and now test
different values for the mutation rate, the recombination rate, the
generation time, the fitness effects, and the population size. To
accomodate this high dimensional data, we have changed figure 4. The
figure now shows the joint distribution of the deleterious effect and
the fraction of synonymous mutations that are deleterious that are
compatible with observations as a heat map. The marginal distribution of
the individual parameters consistent with the data are shown in the supplement.

The distribution of allowed deleterious effects is again peaked around
0.002 and the fraction of "synonymous" mutations that can be neutral is
low, as reported in the earlier version of our manuscript.

##############################################################################
>3) The reasonableness of the ranges for the various parameters should be argued
>with citation to existing literature.

We have added a discussion of the parameter ranges to the methods
section.

##############################################################################
>4) I am a bit surprised that the authors chose to do random sampling in their
>multidimensional parameter space.  One would have expected a Latin hypercube
>sampling to provide a better coverage.

Indeed. Our previous simulations were not limited by coverage of the
parameter space. Now that we cover a much larger (higher dimensional)
space, we did implement a Latin hypercube sampling scheme. We thank the
reviewer for this very useful suggestion.

##############################################################################
>5) The actual dynamics of HIV in the body is presumably more complex: it is
>likely that the virions are produced in large bursts of virtually genetically
>identical virions most of which are then cleared by the body fast on the
>time-scale of HIV generations. There is no discussion whether this leads to
>increased stochasticity in the dynamics, nor whether this can influence the
>quantitative evaluations presented here. It is possible that the authors mean
>to absorb these into some kind of reduced effective population size, but a
>short discussion would be beneficial.

We agree with reviewer that HIV evolution is bursty, but we would like
to mention that actual number of cells that get infected every day is
around 10^7 and hence much larger than we are able to simulate. The
population size in our simulation can therefore be viewed as an
effective population size. We have investigated the sensitivity of our
results to changes in the population size and find that our conclusions
are insensitive to the population size as long as the product of
population size and mutation rate is not very much smaller than 1.

We have added a discussion of our choice of population size to the
methods section next the part where the other parameter choices are
discussed.

##############################################################################
>6) In the discussions, the rise of synonymous mutations is attributed to draft,
>but the fall to selective effects. In general, some of the fall might also be
>due to overlapping selective sweeps: the manuscript does not discuss this
>possibility.

Yes, much of the stochastic component in the dynamics of the frequencies
of synonymous mutations will be due to draft. However, the consistent
loss of variants, conditional on them being at high frequency earlier,
implies selection against these variants. Without a deleterious effect,
the mutations would fix with a probability equal to their frequency,
regardless of whether they are subject to drift of draft. We have added
a sentence to the discussion to make this point clear.

##############################################################################
>7) The observed neutral fixation pattern of the nonsynonymous mutations is
>mentioned, and two explanations have been proposed to duet of increased
>diversity over synonymous mutations with no increase of fixation over neutral
>expectations.  Neither of the explanations, however, are plausible (at least,
>without argument) absent experimental evidence not mentioned in the text. It
>would perhaps be better to either strengthen this discussion with citation to
>literature, or to weaken it by mentioning these as mere possibilities amongst
>others.

We agree with the reviewer that the both hypothesis we put forward to
explain the fixation pattern of non-synonymous mutations are somewhat
tentative. Qualitatively, however, the processes we discuss have been
observed. We now cite additional experimental evidence (Wei et al,
2003). It is common for escape to happen in multiple equivalent ways and
this has been well documented fro early CTL escape and early nAb escape
(Wei et al, Nature, 2003. Bar et al, PPath, 2012). Data for chronic
infection is less conclusive and selection pressures are generally
smaller, but turnover at targeted epitopes and the replacement of
variants that where previously at high frequencies was observed 3y into
the infection in Wei et al, Nature, 2003.  Whether this fall in
frequency of some non-synonymous variants happens because they are
replaced by a less costly escape variant or because the escape variant
itself is targeted is hard to disentangle. We think it is reasonable
that both of these effects contribute. Whether they are sufficient to
produce the observed fixation patterns of non-synonymous mutations is
largely an empirical question. We simply explored using numerical
experiments whether these effects push the curve in the "right"
direction. We have added a disclaimer that more data is needed to fully
explain the pattern of fixation observed for non-synonymous mutations.